---
title: "final DATA SCIENCE"
author: "Zhipeng Chen"
date: "2022/5/4"
output: html_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(mdsr)
library(readr)
library(ggplot2)

options(scipen = 999)
```

# importing data
```{r}
Happy2020 <- read_csv("2020.csv")

Happy2021 <- read_csv("2021.csv")

covidData <- read_csv("worldometer_coronavirus_summary_data.csv")
View(worldometer_coronavirus_summary_data)

summary(Happy2020)
str(Happy2020)
```




# data pre-wrangling 
```{r}

Happy2021 <- Happy2021 %>%
    arrange(desc(`Ladder score`))
```




# making horizontal bar chart for data visualization

1. reorder data presented in the bar chart
2. coord_flip
3. get top 50 countries

```{r}
ggplot(data = Happy2020[1:50, ], aes(x = reorder(`Country name`, `Ladder score`), y = `Ladder score`)) +
    geom_bar(stat='identity', position = "dodge", fill = "light blue") +
    coord_flip() +
    geom_text(aes(label = `Ladder score`), size = 4, hjust = -0.2) +
    labs(x = "country name",
         title = "Top 50 happinest countries in the world from 2019 to 2020") +
    scale_y_continuous(limits = c(0,10)) +
    scale_x_discrete() +
    theme(plot.title = element_text(hjust = 4, face = "bold", size = 15),
          plot.background = element_rect(fill = "#F0F0F0"),
          panel.background = element_rect(color = "#F0F0F0", fill = "#F0F0F0"),
          axis.ticks=element_line(size = 1, color = "#AFB3B3"),
          axis.ticks.length=unit(.25, "cm"),
          axis.text.y = element_text(vjust = 0.000001))
```

# correlations

```{r}

pairs(Happy2020[,3:9], main="Scatterplot Matrix of the happiness score Data",lower.panel=NULL)
```
We can see that there is strong positive linear relationship between the happiness score and GDP, social support, healthy life expectancy, and freedom.

```{r full_model}
full.model <- lm(Happy2020$`Ladder score` ~ Happy2020$`Logged GDP per capita`+Happy2020$`Social support`+Happy2020$`Healthy life expectancy`+Happy2020$`Freedom to make life choices`+Happy2020$Generosity+Happy2020$`Perceptions of corruption`)

par(mfrow=c(1,2))
plot(full.model,1,add.smooth=FALSE)
plot(full.model,2)
```


```{r happiness_data}
head(Happy2020, 20)

head(Happy2021, 20)

unique(Happy2020$`Regional indicator`)

unique(Happy2021$`Regional indicator`)
```


# make new datasets(data wrangling)
```{r data_}
summary2020 <- Happy2020 %>%
    group_by(`Regional indicator`) %>%
    summarise(average_score = mean(`Ladder score`),
              average_GDP = mean(`Logged GDP per capita`),
              average_social_support = mean(`Social support`),
              average_expetancy = mean(`Healthy life expectancy`),
              average_freedom = mean(`Freedom to make life choices`),
              average_generous = mean(Generosity),
              average_corruption = mean(`Perceptions of corruption`),
               year = 2020) 

summary2020$year = as.character(summary2020$year)


summary2021 <- Happy2021 %>%
    group_by(`Regional indicator`) %>%
    summarise(average_score = mean(`Ladder score`),
              average_GDP = mean(`Logged GDP per capita`),
              average_social_support = mean(`Social support`),
              average_expetancy = mean(`Healthy life expectancy`),
              average_freedom = mean(`Freedom to make life choices`),
              average_generous = mean(Generosity),
              average_corruption = mean(`Perceptions of corruption`),
              year = 2021)
    
    summary2021$year = as.character(summary2021$year)
    
```



# merging data
```{r}
total <- rbind(summary2020, summary2021)
```


# grouped bar chart

```{r}
ggplot(data = total, aes(group = year, fill= year, y= average_score, x= reorder(total$`Regional indicator`, average_score))) + 
    geom_bar(position="dodge", stat="identity") +
    labs(title = "bar chart for average score from 2020 to 2021",
         x= "region",
         y = "average score") +
    theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 8),
          plot.background = element_rect(fill = "#F0F0F0"),
        panel.background = element_rect(color = "#F0F0F0", fill = "#F0F0F0"),
        legend.background = element_rect(color = "#F0F0F0", fill = "#F0F0F0"),
        legend.box.background = element_rect(color = "#F0F0F0", fill = "#F0F0F0")) +
    scale_y_continuous(limits = c(0,10),breaks =  c(0,2,4,6,8,10))
```


# covid and happiness score

```{r left_join}
left_join <- Happy2021 %>%
    left_join(covidData, by = c("Country name" = "country")) 

str(left_join)
```



```{r}
ggplot(data = left_join, aes(x = `total_cases_per_1m_population`, y = `Ladder score`)) +
    geom_point() +
    geom_smooth()

ggplot(data = left_join, aes(x = `total_deaths_per_1m_population`, y = `Ladder score`)) +
    geom_point() +
    geom_smooth()
```

















